---
output: 
  html_document: 
    fig_height: 8
    fig_width: 10
    theme: flatly
---

## Preface

This demo blog post uses a very developmental version of `fmi`. To run the demo
on your own computer, you will have to do the following:

```{r install, eval=FALSE}
# fmi depends on rwfs
install_github("ropengov/rwfs")
install_github("ropengov/fmi")
```

## Accessing FMI lightning strikes data using package fmi

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
library(fmi)
library(dplyr)
library(DT)
library(leaflet)
library(raster)
library(sp)
library(tmap)
```


 **NOTE!** You will need to provide
[your own apiKey](http://en.ilmatieteenlaitos.fi/open-data-manual-fmi-wfs-services) first:

```{r apiKey-visible}
API_KEY <- "ENTER YOUR API KEY HERE"
```

```{r apiKey-invisible,echo=FALSE}
API_KEY <- '0253711c-056f-439b-af9d-a5a3a2920faf'
```


```{r setup-bg-map, message=FALSE}
# Load data from tmap
data(Europe)
finland <- subset(Europe, iso_a3 == "FIN")
# Project to WGS84 (EPSG:4326) from ETRS-LAEA (EPSG:3035)
crs_wgs84 <- CRS("+init=epsg:4326")
finland <- spTransform(finland, crs_wgs84)
```

Using the function above simplifies querying. For demonstration, let's
get data for April 2015:

```{r data-query-lightning, message=FALSE, warning=FALSE}
# Define the query time period
startDateTime <- "2015-06-24"
endDateTime <- "2015-06-25"

# Setup a bounding box that covers the whole of Finland
bbox <- raster::extent(c(25, 31, 60, 70))

# First, initialize a new fmi session
init_session(API_KEY)

# NOTE: the response SpatialPointsDataFrame will be in WGS84 CRS.
lightning_data <- fmi_lightnings(startDateTime, endDateTime, bbox)
```

Let's take a look at the measurement data:
```{r data-lightning, message=FALSE, warning=FALSE}
DT::datatable(lightning_data@data)
```

```{r plot-map}
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("navy", "red"), domain = c(0, 1))

m <- leaflet::leaflet(lightning_data) %>%
  addCircleMarkers(
    radius = 3,
    color = ~pal(cloud_indicator),
    stroke = FALSE, fillOpacity = 0.5
  ) %>% 
  addLegend("bottomright", pal = pal, values = ~cloud_indicator,
    title = "Cloud lightning", labels = c("ground", "cloud"),
    opacity = 1
  )
m %>% addProviderTiles("CartoDB.Positron") 

```


That's it, a small demo on how to use `fmi` package for fetching lightning 
strike data.

```{r session-info}
sessionInfo()
```
